Estimation of Intracortical Connectivity in a Dynamical Neural Field Model

### 2010-03-04 ltx/BrainIDE.tex
colored noise derivation is added in the appendix
updated ltx doco to include pde

\section{Neural Field Model}
In this section, we describe a variant of the  Wilson and Cowan neural field model~\cite{Wilson1973} that will be used to create our intracortical connectivity estimator. This model is descriptive of a range of neurodynamics of the cortex such as evoked potentials, visual hallucinations, and epileptic behavior. It is also capable of generating complex patterns of activity such as Turing patterns, spirals, and traveling oscillations. The neural field model is popular due to being parsimonious yet having a strong link with the underlying physiology.
The model relates the average number of action potentials $g$ arriving at $r$ to the local post synaptic membrane voltage. The post-synaptic potentials generated at a neuronal population at location $r$ by action potentials arriving from all other connected populations at locations $r'$ can be described by
\begin{equation}\label{SpikesToPotential}
v\left( {r,t} \right) = \int_{ - \infty }^t {h\left( {t - t'} \right)g\left( {r,t'} \right)dt'} + e\left( {r,t} \right).
\end{equation}
The term $e(r,t)$ can be considered the potential generated by unmodeled inputs and spontaneous background activity. The post-synaptic response kernel $h(t)$ is described by
\begin{equation}\label{SynapticRespKernel}
h(t) = \left\{ {\begin{array}{*{20}{c}}
   {\exp ( - \alpha t)} & {t \ge 0}  \\
   0 & {t < 0}  \\
\end{array}} \right.
\end{equation}
where $\alpha=\tau^{-1}$ and $\tau$ is the synaptic time constant. The synaptic response kernel can be more elaborate, using rise and fall time constants, however, this simple form serves our purposes. Nonlocal interactions between cortical populations are described by	
\begin{equation}\label{RateBasedInteractions}
g\left( {r,t} \right) = \int_\Omega  {w\left( {r,r'} \right)f\left( {v\left( {r',\bar t} \right)} \right)dr'}
\end{equation}
where $\bar t = t - d\left| {r - r'} \right|$, $d$ is the propagation delay of action potentials, $f$ is spiking rate of populations $r'$, $w$ is the spatial connectivity kernel, and $\Omega$ is the spatial domain, representing a cortical sheet or surface.

The average firing rate of the presynaptic neurons is related to the postsynaptic membrane potential by the sigmoidal activation function
\begin{equation}\label{ActivationFunction}
f\left( {{v_t}\left( {r'} \right)} \right) = \frac{{{\nu _0}}}{{1 + \exp \left( {\beta \left( {{v_0}\left( r \right) - {v}\left(r',\bar{t} \right)} \right)} \right)}}.
\end{equation}
%\begin{equation}\label{RateOfPreSynPops}
%p\left( {v\left( {r,\bar t} \right)} \right) = {\nu_0}u\left( {v\left( {r,\bar t} \right) - {v_0}\left( r \right)} \right)
%\end{equation}
%\begin{equation}\label{HeavisideDef}
%u\left( {v\left( {r,\bar t} \right) - {v_0}\left( r \right)} \right) = \left\{ {\begin{array}{*{20}{c}}
%   1 & {v\left( {r,\bar t} \right) \ge {v_0}\left( r \right)}  \\
%   0 & {otherwise}  \\
%\end{array}} \right.
%\end{equation}
Even though we will model  $v_0$ as being homogeneous across the cortical sheet we will express it as being spatially dependant allowing for generality when constructing the state-space representation in the next section. The parameters in equation~\ref{ActivationFunction} are detailed in Table 1. By combining equations~\ref{SpikesToPotential} and \ref{RateBasedInteractions} we get the spatiotemporal model
\begin{equation}\label{FullDoubleIntModel}
v\left(r,t\right) = \int\limits_{-\infty}^t\int\limits_\Omega  h\left(t - t'\right)w\left(r,r'\right)f\left( v\left( r',\bar t \right)\right)dr' dt'+e\left(r,t\right)
\end{equation}
To arrive at the final form of the model we shall state the synaptic response kernel as a Green's function
\begin{equation}\label{GreensFuncDef}
Lh\left( t \right) = \delta \left( t \right),
\end{equation}
where $L$ is a temporal differential operator and $\delta(t)$ is the Dirac-delta function. This provides the most general form
\begin{equation}\label{GenForm}
L\left( {\frac{\partial }{{dt}}} \right)v\left( {r,t} \right) = \int_\Omega  {w\left( {r,r'} \right)f\left( {v\left( {r',\bar t} \right)} \right)dr'}  + e\left( {r,t} \right),
\end{equation}
where $L$ is a polynomial of order $n$ with constant coefficients that provides an $n^{th}$ order model. Most studies considering this model use first or second order time derivatives. We shall use the first order model giving
\begin{equation}\label{FinalForm1}
\frac{{dv\left( {r,t} \right)}}{{dt}} + \alpha{v\left( {r,t} \right)} = \int_\Omega  {w\left( {r,r'} \right)f\left( {v\left( {r',\bar t} \right)} \right)dr'}  +e\left( {r,t} \right).
\end{equation}

% \subsection{Multi-Layer Model}
% \begin{equation}\label{ExcitatoryLayer}
% \frac{dv_e\left( {r,t} \right)}{{dt}} + \alpha_ev_e\left( {r,t} \right) = \int_\Omega  {w_{ee}\left( {r,r'} \right)f\left( v_e\left( r',\bar t \right) \right) + w_{ei}\left( {r,r'} \right)f\left( v_i\left( r',\bar t \right) \right)dr'}  +\varepsilon_e\left( {r,t} \right)
% \end{equation}
% \begin{equation}\label{InhibitoryLayer}
% \frac{dv_i\left( {r,t} \right)}{{dt}} + \alpha_iv_i\left( {r,t} \right) = \int_\Omega  {w_{ie}\left( {r,r'} \right)f\left( v_e\left( r',\bar t \right) \right) + w_{ii}\left( {r,r'} \right)f\left( v_i\left( r',\bar t \right) \right)dr'}  +\varepsilon_i\left( {r,t} \right)
% \end{equation}
% The subscript  denotes connectivity from an inhibitory neural element to an excitatory neural element. We can write the model in matrix form
% \begin{equation}\label{matrixEquation}
% \left[ {\begin{array}{*{20}{c}}
%    {\frac{{d{v_e}\left( t \right)}}
% {{dt}}}  \
%    {\frac{{d{v_i}\left( t \right)}}
% {{dt}}}  \
% 
%  \end{array} } \right] = \int_\Omega  {\left[ {\begin{array}{*{20}{c}}
%    {{w_{ee}}\left( {r,r'} \right)} & {{w_{ei}}\left( {r,r'} \right)}  \
%    {{w_{ie}}\left( {r,r'} \right)} & {{w_{ii}}\left( {r,r'} \right)}  \
% 
%  \end{array} } \right]} \left[ {\begin{array}{*{20}{c}}
%    {f\left( {{v_e}\left( {r',\bar{t}} \right)} \right)}  \
%    {f\left( {{v_i}\left( {r',\bar{t}} \right)} \right)}  \
% 
%  \end{array} } \right]dr' - \left[ {\begin{array}{*{20}{c}}
%    {{\alpha _e}{v_e}\left( {r,t} \right)}  \
%    {{\alpha _i}{v_i}\left( {r,t} \right)}  \
% 
%  \end{array} } \right] + \left[ {\begin{array}{*{20}{c}}
%    {{\varepsilon _e}\left( t \right)}  \
%    {{\varepsilon _i}\left( t \right)}  \
% 
%  \end{array} } \right]
% \end{equation}
% 
% \begin{equation}\label{matrix_equation_in_vector_notation}
%    \frac{d\mathbf{V}(r,t)}{dt} = \int_\Omega \mathbf{W}(r,r')f\left(\mathbf{V}(r,t)\right)dr' - \mathbf{A}\mathbf{V}(r,t) + \mathbf{E}(r,t)
% \end{equation}
% 
% The iEEG observation equation is
% \begin{equation}\label{iEEGObservationEquation}
%     y(k,t) = \int_{\Omega} m(k,r)\left(v_e(r,t) + v_i(r,t)\right)dr' + w(k,t)
% \end{equation}{}

@@

\section{State-Space Model}
We begin the derivation of the state-space model with the discrete time, continuous space IDE neural field model. A one-step Euler method was used to transform the ODE to a discrete time form.
\begin{equation}\label{SystemEq}
\frac{v_{t + 1}\left( \mathbf{r} \right) - v_t(\mathbf{r})}{T_s} = \int_\Omega  {k\left( {{\bf{r}} , {\bf{r'}}} \right)f\left( {{v_t}\left( \bf{r'} \right)} \right)d{\bf{r'}}}  - \alpha {v_t}\left( {\bf{r}} \right) + e_t\left( {\mathbf{r}} \right),
\end{equation}
\begin{equation}\label{SystemEq}
v_{t + 1}\left( \mathbf{r} \right)  = T_s\int_\Omega  {k\left( {{\bf{r}} , {\bf{r'}}} \right)f\left( {{v_t}\left( \bf{r'} \right)} \right)d{\bf{r'}}} - T_s\alpha {v_t}\left( {\bf{r}} \right) + v_t(\mathbf{r}) + T_se_t\left( {\mathbf{r}} \right),
\end{equation}
\begin{equation}\label{SystemEq}
v_{t + 1}\left( \mathbf{r} \right)  = T_s\int_\Omega  {k\left( {{\bf{r}} , {\bf{r'}}} \right)f\left( {{v_t}\left( \mathbf{r'} \right)} \right)d{\bf{r'}}} + (1 - T_s\alpha) v_t\left( {\bf{r}} \right) + T_se_t\left( {\mathbf{r}} \right),
\end{equation}
\begin{equation}\label{SystemEq}
v_{t + 1}\left( \mathbf{r} \right)  = T_s\int_\Omega  {k\left( {{\bf{r}} , {\bf{r'}}} \right)f\left( {{v_t}\left( \bf{r'} \right)} \right)d{\bf{r'}}} + \lambda v_t\left( {\bf{r}} \right) + T_se_t\left( {\mathbf{r}} \right).
\end{equation}
where $e_t(r)\sim \mathcal{GP}(0,R)$, $\lambda = 1-T_s\alpha$ and $T_s$ is the time step or sampling step.

@@

% you can choose not to have a title for an appendix
% if you want by leaving the argument blank
\section{Parameter estimation using LS}
\begin{equation}
 \mathbf x_{t+1}=\mathbf x_{t}+T_s \mathbf Q(\mathbf x_t)\mathbf \theta-\alpha T_s\mathbf x_t+T_s\boldsymbol\Gamma^{-1}\int_\Omega\boldsymbol\phi^\top(\mathbf r)\varepsilon_t(\mathbf r)d\mathbf r
\label{eq:statespacemodel}
\end{equation}
where the matrix $\mathbf Q_{n_x \times n_{\theta}}(\mathbf x_t)$ is defined as
\begin{equation}
 Q_{n_x \times n_{\theta}}=\boldsymbol\Gamma^{-1}\int_\Omega\boldsymbol\phi(\mathbf r)\int_\Omega f(\boldsymbol\phi^\top(\mathbf r')\mathbf x_t)\boldsymbol\psi^\top(\mathbf r-\mathbf r')d\mathbf r'd\mathbf r
\end{equation}
and $T_s$ is the sampling period. Least Square method can be used as (\ref{eq:statespacemodel}) is linear in parameters. At each time instant we have

>>>>

\section{Neural Field Model}
In this section, we describe a variant of the  Wilson and Cowan neural field model~\cite{Wilson1973} that will be used to create our intracortical connectivity estimator. This model is descriptive of a range of neurodynamics of the cortex such as evoked potentials, visual hallucinations, and epileptic behavior. It is also capable of generating complex patterns of activity such as Turing patterns, spirals, and traveling oscillations. The neural field model is popular due to being parsimonious yet having a strong link with the underlying physiology.
The model relates the average number of action potentials $g$ arriving at $r$ to the local post synaptic membrane voltage. The post-synaptic potentials generated at a neuronal population at location $r$ by action potentials arriving from all other connected populations at locations $r'$ can be described by
\begin{equation}\label{SpikesToPotential}
v\left( {r,t} \right) = \int_{ - \infty }^t {h\left( {t - t'} \right)g\left( {r,t'} \right)dt'} + de\left( {r,t} \right).
\end{equation}
The term $e(r,t)$ can be considered the potential generated by unmodeled inputs and spontaneous background activity and is a zero mean temporal Wiener process with incremental covariance $\Sigma_e(r)dt$. The post-synaptic response kernel $h(t)$ is described by
\begin{equation}\label{SynapticRespKernel}
h(t) = \left\{ {\begin{array}{*{20}{c}}
   {\exp ( - \alpha t)} & {t \ge 0}  \\
   0 & {t < 0}  \\
\end{array}} \right.
\end{equation}
where $\alpha=\tau^{-1}$ and $\tau$ is the synaptic time constant. The synaptic response kernel can be more elaborate, using rise and fall time constants, however, this simple form serves our purposes. Nonlocal interactions between cortical populations are described by	
\begin{equation}\label{RateBasedInteractions}
g\left( r,t \right) = \int_\Omega  {w\left( r,r' \right)f\left( v\left( r',\bar t \right) \right)dr'}
\end{equation}
where $\bar t = t - d\left| {r - r'} \right|$, $d$ is the propagation delay of action potentials, $f$ is spiking rate of populations $r'$, $w$ is the spatial connectivity kernel, and $\Omega$ is the spatial domain, representing a cortical sheet or surface.

The average firing rate of the presynaptic neurons is related to the postsynaptic membrane potential by the sigmoidal activation function
\begin{equation}\label{ActivationFunction}
f\left( {{v_t}\left( {r'} \right)} \right) = \frac{{{\nu _0}}}{{1 + \exp \left( {\beta \left( {{v_0}\left( r \right) - {v}\left(r',\bar{t} \right)} \right)} \right)}}.
\end{equation}
%\begin{equation}\label{RateOfPreSynPops}
%p\left( {v\left( {r,\bar t} \right)} \right) = {\nu_0}u\left( {v\left( {r,\bar t} \right) - {v_0}\left( r \right)} \right)
%\begin{equation}\label{HeavisideDef}
%u\left( {v\left( {r,\bar t} \right) - {v_0}\left( r \right)} \right) = \left\{ {\begin{array}{*{20}{c}}
%   1 & {v\left( {r,\bar t} \right) \ge {v_0}\left( r \right)}  \\
%   0 & {otherwise}  \\
%\end{array}} \right.
Even though we will model  $v_0$ as being homogeneous across the cortical sheet we will express it as being spatially dependant allowing for generality when constructing the state-space representation in the next section. The parameters in equation~\ref{ActivationFunction} are detailed in Table 1. By combining equations~\ref{SpikesToPotential} and \ref{RateBasedInteractions} we get the spatiotemporal model
\begin{equation}\label{FullDoubleIntModel}
v\left(r,t\right) = \int\limits_{-\infty}^t\int\limits_\Omega  h\left(t - t'\right)w\left(r,r'\right)f\left( v\left( r',\bar t \right)\right)dr' dt'+de\left(r,t\right)
\end{equation}
To arrive at the final form of the model we shall state the synaptic response kernel as a Green's function
\begin{equation}\label{GreensFuncDef}
Lh\left( t \right) = \delta \left( t \right),
\end{equation}
where $L$ is a temporal differential operator and $\delta(t)$ is the Dirac-delta function. This provides the most general form
\begin{equation}\label{GenForm}
L\left( \frac{\partial }{dt} \right)v\left( {r,t} \right) = \int_\Omega  {w\left( {r,r'} \right)f\left( {v\left( {r',\bar t} \right)} \right)dr'}  + de\left( r,t \right),
\end{equation}
where $L$ is a polynomial of order $n$ with constant coefficients that provides an $n^{th}$ order model. Most studies considering this model use first or second order time derivatives. We shall use the first order model giving
\begin{equation}\label{FinalForm1}
\frac{dv\left( {r,t} \right)}{dt} + \alpha v\left( r,t \right) = \int_\Omega  {w\left( r,r' \right)f\left( {v\left( r',\bar t \right)} \right)dr'}  +de\left( r,t \right).
\end{equation}
The PDE form of the model is
\begin{equation}\label{PDE}
\frac{\partial^2}{\partial t \partial r'} v\left( r,t \right) + \alpha \frac{\partial}{\partial r'} v\left( r,t \right) = w\left( r,r' \right)f\left( v\left( r',\bar t \right) \right)  + \frac{\partial}{\partial r'}de\left( {r,t} \right).
\end{equation}
% \subsection{Multi-Layer Model}
% \begin{equation}\label{ExcitatoryLayer}
% \frac{dv_e\left( {r,t} \right)}{{dt}} + \alpha_ev_e\left( {r,t} \right) = \int_\Omega  {w_{ee}\left( {r,r'} \right)f\left( v_e\left( r',\bar t \right) \right) + w_{ei}\left( {r,r'} \right)f\left( v_i\left( r',\bar t \right) \right)dr'}  +\varepsilon_e\left( {r,t} \right)
% \end{equation}
% \begin{equation}\label{InhibitoryLayer}
% \frac{dv_i\left( {r,t} \right)}{{dt}} + \alpha_iv_i\left( {r,t} \right) = \int_\Omega  {w_{ie}\left( {r,r'} \right)f\left( v_e\left( r',\bar t \right) \right) + w_{ii}\left( {r,r'} \right)f\left( v_i\left( r',\bar t \right) \right)dr'}  +\varepsilon_i\left( {r,t} \right)
% \end{equation}
% The subscript  denotes connectivity from an inhibitory neural element to an excitatory neural element. We can write the model in matrix form
% \begin{equation}\label{matrixEquation}
% \left[ {\begin{array}{*{20}{c}}
%    {\frac{{d{v_e}\left( t \right)}}
% {{dt}}}  \
%    {\frac{{d{v_i}\left( t \right)}}
% {{dt}}}  \
%  \end{array} } \right] = \int_\Omega  {\left[ {\begin{array}{*{20}{c}}
%    {{w_{ee}}\left( {r,r'} \right)} & {{w_{ei}}\left( {r,r'} \right)}  \
%    {{w_{ie}}\left( {r,r'} \right)} & {{w_{ii}}\left( {r,r'} \right)}  \
%  \end{array} } \right]} \left[ {\begin{array}{*{20}{c}}
%    {f\left( {{v_e}\left( {r',\bar{t}} \right)} \right)}  \
%    {f\left( {{v_i}\left( {r',\bar{t}} \right)} \right)}  \
%  \end{array} } \right]dr' - \left[ {\begin{array}{*{20}{c}}
%    {{\alpha _e}{v_e}\left( {r,t} \right)}  \
%    {{\alpha _i}{v_i}\left( {r,t} \right)}  \
%  \end{array} } \right] + \left[ {\begin{array}{*{20}{c}}
%    {{\varepsilon _e}\left( t \right)}  \
%    {{\varepsilon _i}\left( t \right)}  \
%  \end{array} } \right]
% \end{equation}
% \begin{equation}\label{matrix_equation_in_vector_notation}
%    \frac{d\mathbf{V}(r,t)}{dt} = \int_\Omega \mathbf{W}(r,r')f\left(\mathbf{V}(r,t)\right)dr' - \mathbf{A}\mathbf{V}(r,t) + \mathbf{E}(r,t)
% \end{equation}
% The iEEG observation equation is
% \begin{equation}\label{iEEGObservationEquation}
%     y(k,t) = \int_{\Omega} m(k,r)\left(v_e(r,t) + v_i(r,t)\right)dr' + w(k,t)
% \end{equation}{}

@@

\section{State-Space Model}
We begin the derivation of the state-space model with the discrete time, continuous space IDE neural field model. A one-step Euler method was used to transform the nonlinear stochastic IDE to a discrete time form.
\begin{equation}\label{SystemEq}
\frac{v_{t + 1}\left( \mathbf{r} \right) - v_t(\mathbf{r})}{T_s} = \int_\Omega  {k\left( \mathbf{r} , \mathbf{r'} \right)f\left( {v_t\left( \mathbf{r'} \right)} \right)d{\bf{r'}}}  - \alpha v_t\left( \mathbf{r} \right) + e_t\left( \mathbf{r} \right),
\end{equation}
\begin{equation}\label{SystemEq}
v_{t + 1}\left( \mathbf{r} \right)  = T_s\int_\Omega  {k\left( {{\bf{r}} , {\bf{r'}}} \right)f\left( {{v_t}\left( \bf{r'} \right)} \right)d{\bf{r'}}} - T_s\alpha {v_t}\left( {\bf{r}} \right) + v_t(\mathbf{r}) + T_se_t\left( {\mathbf{r}} \right),
\end{equation}
\begin{equation}\label{SystemEq}
v_{t + 1}\left( \mathbf{r} \right)  = T_s\int_\Omega  {k\left( {{\bf{r}} , {\bf{r'}}} \right)f\left( {{v_t}\left( \mathbf{r'} \right)} \right)d{\bf{r'}}} + (1 - T_s\alpha) v_t\left( {\bf{r}} \right) + T_se_t\left( {\mathbf{r}} \right),
\end{equation}
\begin{equation}\label{SystemEq}
v_{t + 1}\left( \mathbf{r} \right)  = T_s\int_\Omega  {k\left( {{\bf{r}} , {\bf{r'}}} \right)f\left( {{v_t}\left( \bf{r'} \right)} \right)d{\bf{r'}}} + \lambda v_t\left( {\bf{r}} \right) + T_se_t\left( {\mathbf{r}} \right).
\end{equation}
where $e_t(r)\sim \mathcal{GP}(0,R)$, $\lambda = 1-T_s\alpha$ and $T_s$ is the time step or sampling step.

@@

% you can choose not to have a title for an appendix
% if you want by leaving the argument blank
\section{Colored noise}
We have
\begin{equation}
v_{t + 1}\left( \mathbf{r} \right)  = T_s\int_\Omega  {k\left( {{\bf{r}} , {\bf{r'}}} \right)f\left( {{v_t}\left( \bf{r'} \right)} \right)d{\bf{r'}}} + \lambda v_t\left( {\bf{r}} \right) + T_se_t\left( {\mathbf{r}} \right).
\end{equation}
where the disturbance $e_t(\mathbf r)$ is a zero-mean normally distributed noise process, spatially coloured but temporally independent, with covariance

\begin{equation}
cov(e_t(\mathbf{s}),e_{t+\tau}(\mathbf{r}))=
\begin{cases}
\beta\left(\mathbf{s}-\mathbf{r}\right), & \tau=0 \\
0 & \mathrm{otherwise}
\end{cases}
\label{eq:FieldCovariance}
\end{equation}
The final state-space model is given by
\begin{equation}
x_{t + 1} = T_sq(x_t)\theta -\lambda x_t + T_s\Gamma^{-1}\int_\Omega  {\phi ( \mathbf{r} )e_t( \mathbf{r} )d\mathbf{r}}
\end{equation}
then the covariance of the noise term is given by
\begin{eqnarray}
\mathbf \Sigma_w&{}={}&T_s^2 \times \mathbf{\Gamma}^{-1}\mathbf E[\int_{\mathcal{S}}\boldsymbol\phi\left(\mathbf s\right)e_t\left(\mathbf s\right)d\mathbf s\int_{\mathcal{S}}\boldsymbol\phi\left(\mathbf r\right)^{\top}e_t\left(\mathbf r\right)d\mathbf r]\mathbf{\Gamma}_{x}^{- \top} \nonumber \\
&=&T_s^2 \times\mathbf{\Gamma}^{-1}\iint\limits_{\mathcal{S}}\boldsymbol\phi\left(\mathbf s\right) \mathbf E[e_t\left(\mathbf s\right)e_t\left(\mathbf r\right)]\boldsymbol\phi\left(\mathbf r\right)^{\top}d\mathbf r d\mathbf s\mathbf{\Gamma}^{- \top} \nonumber\\
&=&{}T_s^2 \times\mathbf{\Gamma}^{-1}\iint\limits_{\mathcal{S}}\boldsymbol\phi\left(\mathbf s\right) \beta\left(\mathbf s- \mathbf r \right)\boldsymbol\phi\left(\mathbf r\right)^{\top}d\mathbf r d\mathbf s\mathbf{\Gamma}^{- \top} 
\end{eqnarray}

\section{Parameter estimation using LS}
\begin{equation}
 \mathbf x_{t+1}=\mathbf x_{t}+T_s \mathbf Q(\mathbf x_t)\mathbf \theta-\alpha T_s\mathbf x_t+T_s\boldsymbol\Gamma^{-1}\int_\Omega\boldsymbol\phi(\mathbf r)\varepsilon_t(\mathbf r)d\mathbf r
\label{eq:statespacemodel}
\end{equation}
where the matrix $\mathbf Q_{n_x \times n_{\theta}}(\mathbf x_t)$ is defined as
\begin{equation}
 Q_{n_x \times n_{\theta}}=\boldsymbol\Gamma^{-1}\int_\Omega\boldsymbol\phi(\mathbf r)\int_\Omega f(\boldsymbol\phi^\top(\mathbf r')\mathbf x_t)\boldsymbol\psi^\top(\mathbf r-\mathbf r')d\mathbf r'd\mathbf r
\end{equation}
and $T_s$ is the sampling period. Least Square method can be used as (\ref{eq:statespacemodel}) is linear in parameters. At each time instant we have
